\subsection{System Testing}

\begin{Frame}{System Testing}
  \begin{itemize}
    \item Difficult, no (agreed) methodology.
    \item Divided into several categories:\newline
      \ \ -- Function testing, \newline
      \ \ -- volume and stress testing, \newline
      \ \ -- usability testing, \newline
      \ \ -- security testing, \newline
      \ \ -- performance testing, \newline
      \ \ -- installability testing and \newline
      \ \ -- recovery testing.
  \end{itemize}
\end{Frame}

\begin{Frame}{Function Testing}
  \begin{itemize}
    \item Does the system fulfill the functional specification?
    \item Create input data for the whole application and check \newline
      \ \ -- its output or\newline
      \ \ -- its behavior.
    \item Do not consider implementation details.
  \end{itemize}
\end{Frame}

\begin{Frame}{Volume and Stress Testing}
  \begin{itemize}
    \item Does the system work correctly in limit situations under \newline
      \ \ -- large volumes of data or\newline
      \ \ -- under high peak volumes of data?
    \item Limit situations: main memory or hard disk almost fully used.
  \end{itemize}
\end{Frame}

\begin{Frame}{Usability Testing}
  \begin{itemize}
    \item Are the error messages useful and understandable?
    \item Is the online-documentation adequate?
    \item Is the user interface adequate for different user types (novice, expert)?
    \item \ldots
  \end{itemize}
\end{Frame}

\begin{Frame}{Security Testing}
  \begin{itemize}
    \item Are access rights correctly implemented?
    \item Are transactions still correct under parallel access?
    \item \ldots
  \end{itemize}
\end{Frame}

\begin{Frame}{Performance Testing}
  \begin{itemize}
    \item Does the system meet the performance requirements?
    \item Possible requirements:\newline
      \ \ -- Transactions per second,\newline
      \ \ -- response time,\newline
      \ \ -- \ldots
  \end{itemize}
\end{Frame}
  
\begin{Frame}{Installability Testing}
  \begin{itemize}
    \item Can the system be installed in all
required environments?
    \item Are required dependencies available in all required environments?
    \item Is it compatible with other software?
    \item \ldots
  \end{itemize}
\end{Frame}

\begin{Frame}{Recovery Testing}
  \begin{itemize}
    \item Does the system recover well from a power cut?
    \item How much data is lost after the recovery?
    \item \ldots
  \end{itemize}
\end{Frame}

