\chapter{\LaTeX-Demo}

\targets{
    \item frames with and without title in the script version
    \item blocks with and without title in the script version
    \item lists and enumerations becoming sentences in the script version
    \item colored blocks and tables
}

\section{frames}

\begin{frame}{I'm a frame}{with subtitle}
  Title and subtitle of this frame are not displayed in the script version.
\end{frame}

\begin{Frame}{Important frame title}{with subtitle}
  This frame will be displayed with title and subtitle in the script version.
\end{Frame}

\subsection{fragile content}

\begin{Frame}[fragile]{fragile frame}{with title in the script version}
  This frame will be displayed with title and subtitle in the script version.
  
  \begin{verbatim}
    And it contains fragile content.
  \end{verbatim}
\end{Frame}

\section{blocks}

\begin{frame}{blocks}
  \begin{block}{title}
    This block will be displayed as normal paragraph in the secript version.
  \end{block}
  
  \begin{Block}{important title}
    This block will be displayed with title in the script version.
  \end{Block}
\end{frame}

\section{lists and enumerations}

\begin{frame}{lists}
  This list
  \begin{beameritemize}
    \item will appear
    \item as a normal sentence
  \end{beameritemize}
  in the script version.
\end{frame}

\begin{frame}{enumerations}
  This enumeration
  \begin{beamerenumerate}
    \item will appear
    \item as normal sentence
    \item still containing the numbering 
  \end{beamerenumerate}
  in the script version.
\end{frame}

\section{colors}

\subsection{blocks}

\begin{frame}{colored blocks}
  \begin{block}{normal block}
    This is a normal block.
  \end{block}
  
  \begin{alertblock}{alertblock}
    This is an alertblock.
  \end{alertblock}
  
  \begin{exampleblock}{exampleblock}
    This is an exampleblock.
  \end{exampleblock}
\end{frame}

\subsection{tables}

\begin{frame}{colored tables}
  \begin{tabular}{ll}
    header & row \\
    This & is a \\
    normal & table
  \end{tabular}
  
  \vskip2ex
  
  \begin{zebratabular}{ll}
    \headerrow header & row \\
    This & is a \\
    normal & table
  \end{zebratabular}
\end{frame}