\section{Principles of Testing}

\subsection*{Motivating Example}

\subsubsection*{Problem Definition}

\begin{Frame}{A little exercise (Myers)}
  \begin{itemize}
    \item Input: three integer values\\
      representing lengths of the sides of a triangle.
    \item Output: \texttt{scalene}, \texttt{isosceles}, or \texttt{equilateral}
  \end{itemize}

  \xxx

  \hfill
  \shortstack{
    \tikz[auto,scale=.5]{
      \draw[draw=maincolor, thick]
        (0,0) -- node {4} ++(-4,0) -- node{3} ++(0,3) coordinate (last) -- cycle;
      \path (last) -- node{5} (0,0);}\\
    \texttt{scalene}\strut\\
    \scriptsize (unregelmäßig)
  }
  \hfill
  \shortstack{
    \tikz[auto,scale=.5]{
      \draw[draw=maincolor, thick]
        (0,0) -- node {3} ++(2,2.2361) -- node{3} ++(2,-2.2361) coordinate (last) -- cycle;
      \path (last) -- node{4} (0,0);}\\
    \texttt{isosceles}\strut\\
    \scriptsize (gleichschenklig)
  }
  \hfill
  \shortstack{
    \tikz[auto,scale=.5]{
      \draw[draw=maincolor, thick]
        (0,0) -- node {4} ++(2,3.4641) -- node{4} ++(2,-3.4641) coordinate (last) -- cycle;
      \path (last) -- node{4} (0,0);}\\
    \texttt{equilateral}\strut\\
    \scriptsize (gleichseitig)
  }
  \hfill\strut

  \xxx

  \begin{Block}{Exercise (10 min)}
    Write a set of \alert{test cases} 
    (concrete values for input and output) that
    you feel would adequately test this program.
  \end{Block}
\end{Frame}

\subsubsection*{Test Cases}

\begin{Frame}[allowframebreaks]{You get one point for each success}{14 possible points}
  Do you have?
  \begin{itemize}
    \item a test case with a valid scalene triangle?\\
      (Note: 1,2,3 and 2,5,10 are no triangles!)
    \item a test case with a valid equilateral triangle
    \item a test case with a valid isosceles triangle\\
      (Note: 2,2,4 is no triangle!)
    \item tried all three permutations of the last case?\\
    (e.g., 3,3,4; 3,4,3 and 4,3,3)
    \item a test case with one side being zero?
    \item a test case with one side being negative?
    \item specified the expected output for each test case?
  \end{itemize}

  \framebreak

  Do you have?
  \begin{itemize}
    \item a test case with $a,b,c > 0$, s.t. $a+b=c$?
    \item tried all three positions for $c$ in the last case?\\
    (e.g., 1,2,3; 1,3,2 and 3,1,2)
    \item a test case with $a,b,c > 0$, s.t. $a+b<c$?
    \item tried all three positions for $c$ in the last case?\\
    (e.g., 1,2,4; 1,4,2; and 4,1,2)
    \item a test case with all sides being zero?
    \item a test case specifying non-integer values?
    \item a test case specifying the wrong number of values?
  \end{itemize}
\end{Frame}
